gss marriage politics analysis

### 2023-04-20 outputs/paper/paper.qmd
added graphs and tables to data section

@@

This paper uses the R statistical programming language [@citeR], along with several packages, tidyverse [@citetidyverse], janitor [@citejanitor], here [@citehere], and [@citedplyr]. All figures in this paper were created using the packages ggplot2 [@citeggplot2] and the tables were created using knitr [@citeknitr] and kableExtra[@citekablextra]. Combinations of figures were created using [@citepatchwork] The color styling of graphs has been created using RColorBrewer \[@citecolorbrewer\]. The model in this paper was created using the rstanarm [@citerstanarm] package.

@@

variable_table <-
  tibble(
    "Variable" = c(
      "partyid",
      "polviews",
      "marital",
      "sex",
      "cohort"
    ),
    "Variable Description" = c(
      "Self-ascribed belonging to the Republican, Democratic, or Indepedent parties",
    "Self-ascribed placement on a Likert scale from 'extremely liberal to extremely conservative'",
    "Marital status of respondent",
    "Self-ascribed sex of respondent",
    "Generational cohort that respondent belongs to "
    )
  )

@@

@@

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

>>>>

#### Ordering ####

cleaned_gss_2021$partyid <- 
  factor(cleaned_gss_2021$partyid, 
         levels = c("Strong Democrat", "Not Very Strong Democrat", "Independent (Close to Democrat)", "Independent (Neither)", "Independent (Close to Republican)", "Not Very Strong Republican", "Strong Republican"))

cleaned_gss_2018$partyid <- 
  factor(cleaned_gss_2018$partyid, 
         levels = c("Strong Democrat", "Not Very Strong Democrat", "Independent (Close to Democrat)", "Independent (Neither)", "Independent (Close to Republican)", "Not Very Strong Republican", "Strong Republican"))

cleaned_gss_2016$partyid <- 
  factor(cleaned_gss_2016$partyid, 
         levels = c("Strong Democrat", "Not Very Strong Democrat", "Independent (Close to Democrat)", "Independent (Neither)", "Independent (Close to Republican)", "Not Very Strong Republican", "Strong Republican"))

cleaned_gss_2014$partyid <- 
  factor(cleaned_gss_2014$partyid, 
         levels = c("Strong Democrat", "Not Very Strong Democrat", "Independent (Close to Democrat)", "Independent (Neither)", "Independent (Close to Republican)", "Not Very Strong Republican", "Strong Republican"))
```

#### Ordering ####

cleaned_gss_2021$marital <- 
  factor(cleaned_gss_2021$marital, 
         levels = c("Married", "Never Married", "Separated", "Divorced", "Widowed"))

cleaned_gss_2018$marital <- 
  factor(cleaned_gss_2018$marital, 
         levels = c("Married", "Never Married", "Separated", "Divorced", "Widowed"))

cleaned_gss_2016$marital <- 
  factor(cleaned_gss_2016$marital, 
         levels = c("Married", "Never Married", "Separated", "Divorced", "Widowed"))

cleaned_gss_2014$marital <- 
  factor(cleaned_gss_2014$marital, 
         levels = c("Married", "Never Married", "Separated", "Divorced", "Widowed"))
```

#### Ordering ####

cleaned_gss_2021$cohort <- 
  factor(cleaned_gss_2021$cohort, 
         levels = c("Gen Z", "Millenial", "Gen X", "Boomer", "Post-War"))

cleaned_gss_2018$cohort <- 
  factor(cleaned_gss_2018$cohort, 
         levels = c("Gen Z", "Millenial", "Gen X", "Boomer", "Post-War"))

cleaned_gss_2016$cohort <- 
  factor(cleaned_gss_2016$cohort, 
         levels = c("Gen Z", "Millenial", "Gen X", "Boomer", "Post-War"))

cleaned_gss_2014$cohort <- 
  factor(cleaned_gss_2014$cohort, 
         levels = c("Millenial", "Gen X", "Boomer", "Post-War"))
```

#### Ordering####

cleaned_gss_2021$polviews <- 
  factor(cleaned_gss_2021$polviews, 
         levels = c("Extremely Conservative", "Conservative", "Slightly Conservative", "Moderate", "Slightly Liberal", "Liberal", "Extremely Liberal"))

cleaned_gss_2018$polviews <- 
  factor(cleaned_gss_2018$polviews, 
         levels = c("Extremely Conservative", "Conservative", "Slightly Conservative", "Moderate", "Slightly Liberal", "Liberal", "Extremely Liberal"))

cleaned_gss_2016$polviews <- 
  factor(cleaned_gss_2016$polviews, 
         levels = c("Extremely Conservative", "Conservative", "Slightly Conservative", "Moderate", "Slightly Liberal", "Liberal", "Extremely Liberal"))

cleaned_gss_2014$polviews <- 
  factor(cleaned_gss_2014$polviews, 
         levels = c("Extremely Conservative", "Conservative", "Slightly Conservative", "Moderate", "Slightly Liberal", "Liberal", "Extremely Liberal"))
```

@@

This paper uses the R statistical programming language [@citeR], along with several packages, tidyverse [@citetidyverse], janitor [@citejanitor], here [@citehere], and [@citedplyr]. All figures in this paper were created using the packages ggplot2 [@citeggplot2] and the tables were created using knitr [@citeknitr] and kableExtra[@citekablextra]. Combinations of figures were created using [@citepatchwork] The color styling of graphs has been created using RColorBrewer [@citecolorbrewer]. The model in this paper was created using the rstanarm [@citerstanarm] package.

@@

variable_table <-
  tibble(
    "Variable" = c(
      "partyid",
      "polviews",
      "marital",
      "sex",
      "cohort"
    ),
    "Variable Description" = c(
      "Self-ascribed belonging to the Republican, Democratic, or Indepedent parties",
    "Self-ascribed placement on the political spectrum from 'extremely liberal to extremely conservative'",
    "Marital status of respondent",
    "Self-ascribed sex of respondent",
    "Generational cohort that respondent belongs to "
    )
  )

@@

## Marital Status and Cohort

```{r}
#| include: false
#| message: false
#### Creating Graph ####
cohort_marital_2021 <- ggplot(data=cleaned_gss_2021, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Birth Cohort",
    y = "Number of Respondents",
    fill = "Marital Status",
    caption = "2021 GSS"
  ) +
  scale_fill_brewer(palette="Spectral")+
  theme(legend.position = "none") + #include legend at bottom to ensure readability when using patchwork package
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false
#### Creating Graph ####
cohort_marital_2018 <- ggplot(data=cleaned_gss_2018, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Birth Cohort",
    y = "Number of Respondents ",
    fill = "Marital Status",
    caption = "2018 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") + 
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false

#### Creating Graph####
cohort_marital_2016 <- ggplot(data=cleaned_gss_2016, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Birth Cohort",
    y = "Number of Respondents",
    fill = "Marital Status",
    caption = "2016 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(legend.position = "none")+
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false

#### Creating Graph ####
cohort_marital_2014 <- ggplot(data=cleaned_gss_2014, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Birth Cohort",
    y = "Number of Respondents",
    fill = "Marital Status",
    caption = "2014 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(legend.position = "none") +
  theme(text = element_text(size=6))
```

```{r}
#| echo: false
#| warning: false
#| label: fig-2
#| fig-cap: "2014-2021 Comparisons of Birth Cohort Against Respondent Marital Status"
#### Using Patchwork Package to Combine Two Graphs ####
(cohort_marital_2021) + (cohort_marital_2018) + (cohort_marital_2016) + (cohort_marital_2014) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

## Marital Status and Self-Ascribed Belonging to Political Parties

```{r}
#| include: false
#| message: false

#### Creating Graph ####
partyid_marital_2021 <- ggplot(data=cleaned_gss_2021, aes(x = partyid, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Political Identification",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2021 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none") +
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false

#### Creating Graph ####
partyid_marital_2018 <- ggplot(data=cleaned_gss_2018, aes(x = partyid, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Political Identification",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2018 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none") +
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false

#### Creating Graph ####
partyid_marital_2016 <- ggplot(data=cleaned_gss_2016, aes(x = partyid, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Political Identification",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2016 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none") +
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false

#### Creating Graph ####
partyid_marital_2014 <- ggplot(data=cleaned_gss_2014, aes(x = partyid, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Political Identification",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2014 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(text = element_text(size=6)) 
```

```{r}
#| echo: false
#| warning: false
#| label: fig-3
#| fig-cap: "2014-2021 Comparisons of Birth Cohort Against Respondent Marital Status"
#### Using Patchwork Package to Combine Two Graphs ####
(partyid_marital_2021) + (partyid_marital_2018) + (partyid_marital_2016) + (partyid_marital_2014)+ 
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-4
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2021 Respondents"

cleaned_gss_2021|>
  tabyl(partyid, marital)|>
  kable(col.names = c(
    "Political Identity",
    "Married",
    "Never Married",
    "Separated",
    "Divorced",
    "Widowed"
    ),
    booktabs = TRUE,
    linesep = "")

```{r}
#| echo: false
#| warning: false
#| label: tbl-5
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2018 Respondents"

cleaned_gss_2018|>
  tabyl(partyid, marital)|>
  kable(col.names = c(
    "Political Identity",
    "Married",
    "Never Married",
    "Separated",
    "Divorced",
    "Widowed"
    ),
    booktabs = TRUE,
    linesep = "")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-6
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2016 Respondents"

cleaned_gss_2016|>
  tabyl(partyid, marital)|>
  kable(col.names = c(
    "Political Identity",
    "Married",
    "Never Married",
    "Separated",
    "Divorced",
    "Widowed"
    ),
    booktabs = TRUE,
    linesep = "")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-7
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2014 Respondents"

cleaned_gss_2014|>
  tabyl(partyid, marital)|>
  kable(col.names = c(
    "Political Identity",
    "Married",
    "Never Married",
    "Separated",
    "Divorced",
    "Widowed"
    ),
    booktabs = TRUE,
    linesep = "")
```

## Marital Status and Self-Ascribed Placement on Political Spectrum

```{r}
#| include: false
#| message: false

#### Creating Graph ####
polviews_marital_2021 <- ggplot(data=cleaned_gss_2021, aes(x = polviews, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Political View",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2021 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none") +
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false

#### Creating Graph ####
polviews_marital_2018 <- ggplot(data=cleaned_gss_2018, aes(x = polviews, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Political View",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2018 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none") +
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false

#### Creating Graph ####
polviews_marital_2016 <- ggplot(data=cleaned_gss_2016, aes(x = polviews, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Political View",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2016 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none") +
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false

#### Creating Graph ####
polviews_marital_2014 <- ggplot(data=cleaned_gss_2014, aes(x = polviews, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Political View",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2014 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(text = element_text(size=6))
```

```{r}
#| echo: false
#| warning: false
#| label: fig-4
#| fig-cap: "2014-2021 Comparisons of Birth Cohort Against Respondent Marital Status"
#### Using Patchwork Package to Combine Two Graphs ####
(polviews_marital_2021) + (polviews_marital_2018) + (polviews_marital_2016) + (polviews_marital_2014)+ 
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-8
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2014 Respondents"

cleaned_gss_2021|>
  tabyl(partyid, marital)|>
  kable(col.names = c(
    "Political Identity",
    "Married",
    "Never Married",
    "Separated",
    "Divorced",
    "Widowed"
    ),
    booktabs = TRUE,
    linesep = "")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-9
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2014 Respondents"

cleaned_gss_2018|>
  tabyl(partyid, marital)|>
  kable(col.names = c(
    "Political Identity",
    "Married",
    "Never Married",
    "Separated",
    "Divorced",
    "Widowed"
    ),
    booktabs = TRUE,
    linesep = "")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-10
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2014 Respondents"

cleaned_gss_2016|>
  tabyl(partyid, marital)|>
  kable(col.names = c(
    "Political Identity",
    "Married",
    "Never Married",
    "Separated",
    "Divorced",
    "Widowed"
    ),
    booktabs = TRUE,
    linesep = "")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-11
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2014 Respondents"

cleaned_gss_2014|>
  tabyl(partyid, marital)|>
  kable(col.names = c(
    "Political Identity",
    "Married",
    "Never Married",
    "Separated",
    "Divorced",
    "Widowed"
    ),
    booktabs = TRUE,
    linesep = "")
```

@@
